Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex
first pass at a conclusion

\textbf{Single-action-irreflexive} stencils (3\%) ) consist of a single
\term{backward}, \term{forward}, or \term{centered} irreflexive term
and some number of reflexive terms.
% /home/mrd45/src/corpus/geos-chem/GeosRad/rrtmg_lw_rtrnmc.F90:355
\begin{minted}{fortran}
do lev = nlayers, 1, -1
    ! ...
    radld = radld - radld * (atrans(lev) + &
       efclfrac(lev,igc) * (1. - atrans(lev))) + &
       gassrc + cldfmc(igc,lev) * &
       (bbdtot * atot(lev) - gassrc)
    != stencil readOnce, (forward(depth=1, dim=2, irreflexive)) :: cldfmc
    drad(lev-1) = drad(lev-1) + radld
\end{minted}
Notice that, in this last example, the read from
\mintinline{fortran}{cldfmc} flows through onto the line with the
stencil. Stencils of this type required a neightbourhood index without a zero-offset in one dimension only. The remaining dimensions must either have no offset or not pertain to an induction variable.

@@

\section{Conclusion}

\anote{We need a conclusion!}

>>>>

\textbf{Single-action-irreflexive} stencils (3\%) ) consist of a single
\term{backward}, \term{forward}, or \term{centered} irreflexive region constant
and some number of reflexive region constants.
% /home/mrd45/src/corpus/geos-chem/GeosRad/rrtmg_lw_rtrnmc.F90:355
\begin{minted}{fortran}
do lev = nlayers, 1, -1
    ! ...
    radld = radld - radld * (atrans(lev) + &
       efclfrac(lev,igc) * (1. - atrans(lev))) + &
       gassrc + cldfmc(igc,lev) * &
       (bbdtot * atot(lev) - gassrc)
    != stencil readOnce, (forward(depth=1, dim=2, irreflexive)) :: cldfmc
    drad(lev-1) = drad(lev-1) + radld
\end{minted}
Notice that, in this last example, the read from
\mintinline{fortran}{cldfmc} flows through onto the line with the
stencil. Stencils of this type required a neightbourhood index without a zero-offset in one dimension only. The remaining dimensions must either have no offset or not pertain to an induction variable.

@@

\section{Conclusion and future work}

In this paper we have described an annotation language which can be
used to specify the access pattern of a stencil computation. We
defined a denotational semantic model for our language and proved its
soundness. This gave a basis for an equational theory of
specifications and also justifies our inference and checking
procedure. We evaluated our approach by implementing the stencil inference and
checking procedure in Camfort, an open-source analysis tool\footnote{\url{https://github.com/camfort/camfort}} for
Fortran programs and running the inference process over a corpus of
1.1 million lines of Fortran code.

Our motivation for this work was to provide an intermediate
specification language to allow scientists to verify numerical
computations. The most common types of stencils we found consisted only of
reflexive region constants which are unlikely to give rise to
programming errors. However, around 10\% of stencils covered more
complicated program structures for which indexing errors are much more
plausible. We showed two examples of more complicated stencils for
which CamFort was able to detect a significant majority of possible
indexing errors.

Our definition of a stencil computation meant that we generated no
specification for a large number of iterated array accesses in the corpus. For
future work we will be looking at expanding the definition to allow us
to capture more of these examples.
